---
title: "Preliminary CMF Patent Handlinks"
author: "Terence Chau"
output: html_notebook
---

# Set up

Load patent data up to 1870.

Note: We're going to have to take into account expirations somehow, but ignore for now.

```{r}
library(tidyverse)
library(magrittr)
library(lubridate)
library(haven)
library(stringdist)
library(openxlsx)

patents <- read_csv("~/Dropbox/Work/UChicago/RA/RA Hornbeck/CMF Patents/Terence Working Folders/output/patents_1840_1900.csv") %>% 
  filter(year(disposal_date) <= 1870)
```

Get the distribution of patent counts by counties:

```{r}
inventor_county_names <- patents %>% 
  select(inventor_fips, inventor_county, inventor_state) %>% 
  rename(fips = inventor_fips,
         county = inventor_county,
         state = inventor_state) %>% 
  distinct() 

patent_inventor_county_dist <- table(patents$inventor_fips) %>% 
  data.frame() %>% 
  rename(fips = Var1, 
         n = Freq) %>% 
  arrange(-n) %>% 
  left_join(inventor_county_names, by = "fips") %>% 
  select(fips, state, county, n)

summary(patent_inventor_county_dist$n)
```

Do the same for assignee counties:

```{r}
assignee_county_names <- patents %>% 
  select(assignee_fips, assignee_county, assignee_state) %>% 
  rename(fips = assignee_fips,
         county = assignee_county,
         state = assignee_state) %>% 
  distinct() 

patent_assignee_county_dist <- table(patents$assignee_fips) %>% 
  data.frame() %>% 
  rename(fips = Var1, 
         n = Freq) %>% 
  arrange(-n) %>% 
  left_join(assignee_county_names, by = "fips") %>% 
  select(fips, state, county, n)

summary(patent_assignee_county_dist$n)
```

Inventors and assignees combined:

```{r}
county_names <- bind_rows(inventor_county_names, assignee_county_names) %>% 
  distinct()

patent_county_dist <- patent_inventor_county_dist %>% 
  select(-c(state, county)) %>% 
  rename(inventor_count = n) %>% 
  left_join(patent_assignee_county_dist, by = "fips") %>%
  select(-c(state, county)) %>% 
  rename(assignee_count = n) %>% 
  mutate(total_count = inventor_count + assignee_count,
         total_count = ifelse(is.na(inventor_count), assignee_count, total_count), 
         total_count = ifelse(is.na(assignee_count), inventor_count, total_count)) %>% 
  arrange(-total_count) %>% 
  select(fips, total_count, inventor_count, assignee_count) %>% 
  left_join(county_names, by = "fips") %>% 
  select(fips, state, county, everything()) %>% 
  arrange(-total_count)

summary(patent_county_dist$total_count)
```

```{r}
patent_county_dist
```

# Picking a First County

Whatever patent-establishments links are made at this point will have to 
be redone anyways because the patent county boundaries haven't been fixed, but
might as well look at the pilot counties from the CMF handlinks that have been
deduplicated.

Add up patent counts (both inventor and assignee counts) within the pilot counties:

```{r}
cmf_pilot <- read_csv("~/Dropbox/Work/UChicago/RA/RA Hornbeck/CMF Micro Data/Terence Working Folders/firm_matching/output/crosswalked_counties/1860_to_1870/county_crosswalks/pilot/matching_sample_crosswalk.csv") %>% 
  select(fips_1870, county_1870) %>% 
  rename(fips = fips_1870,
         county = county_1870)

patent_county_dist %>% filter(fips %in% cmf_pilot$fips)
```

It doesn't seem like there was much patenting going on in the pilot CMF linking
sample, so we'll pick a county with ~100 patents, either issued or assigned.
Possible candidates are Newport, RI (44005), Miami, OH (39109), Lehigh, PA (42077),
Lee, IL (17103), Dearborn, IN (18029), Kenton, KY (21117), Racine, WI (55101),
and Sussex, NJ (34037). We can check how many establishments are in each and pick 
accordingly.

Note: Many Illinois counties are missing in the CMF, so we can discard Lee right
away.

```{r}
cmf_1870 <- read_dta("~/Dropbox/Work/UChicago/RA/RA Hornbeck/CMF Micro Data/Vivian Working Folders/industry_assignment/estabs1870.dta") 

cmf_1870 %>%
  filter(fips %in% c(44005, 39109, 42077, 17103, 18029, 21117)) %>% 
  count(fips)
```

Three of these counties are missing CMF data, so let's start with Dearborn.

After handlinking, Dearborn yielded sets of patents linked to about 3 firms, 
2 of which I couldn't tell apart. So maybe including USPC codes in the linking 
sheet would help.

Maybe the problem is that because the patent count distribution is so skewed,
there will be extremely few matches for most counties.

Let's try out a county with more patents.

# Picking a Second County

Looking at counties with around 500 patents, we see that we have way larger
counties. Clark, Ohio (39023) has the smallest establishment to patent ratio 
(184/473), so let's try that one next.

```{r}
cmf_1870 %>%
  filter(fips %in% c(26163, 18097, 36011, 39023, 23005)) %>% 
  count(fips)
```

I found lot more matches in Springfield. It seems like there was a lot more
industrial activity in Springfield, and I think it had some patterns we'll see
in other counties as well. For example, most inventors had many patents to 
their name, and matching rates will basically depend on whether firms tied to
those inventors are in the CMF or not. In fact, a quick search of William N. 
Whiteley and James Leffel reveals that they were lifelong 
engineers/inventors...Leffel built his first
grist mill at age 12 (which made me reassess my life accomplishments to be 
honest) and started designing waterwheels for soon after. He went on to outfit
mills all over the region with his waterwheels then went on to run several
businesses. His company still exists in name. Whiteley had about 42 patents 
under his name by the time he died and had a pretty successful business making 
reapers and other agricultural tools.

One issue I ran into was that when there were several establishments that could 
have been related to these inventors I wasn't able to tell them apart for the
most part. The NBER patent class to industry mappings helped a little, but you
can tell they're geared towards modern data...some of these patents are 
classified as "computers/electronics", which is a stretch.

Here's a summary of the match rates:

```{r}
drop_blank_rows <- function(df){
  df[rowSums(!is.na(df)) > 0, ]
}

dearborn_in <- read_excel("/Users/terencechau/Dropbox/Work/UChicago/RA/RA Hornbeck/CMF Patents/Terence Working Folders/output/pilot_linking_sheets/fips18029_1870.xlsx",
                          col_types = "text") %>% 
  drop_blank_rows()

springfield_oh <- read_excel("/Users/terencechau/Dropbox/Work/UChicago/RA/RA Hornbeck/CMF Patents/Terence Working Folders/output/pilot_linking_sheets/fips39023_1870.xlsx",
                          col_types = "text") %>% 
  drop_blank_rows()

count_estabs_pats <- function(df){
  df %<>% 
    mutate(estab_id = paste(file_name, firm_number, sep = "_"))
  num_estabs <- length(unique(df$estab_id))
  num_pats <- length(unique(df$patent_number))
  
  num_matched_estabs <- df %>% 
    group_by(estab_id) %>% 
    summarize(matched = ifelse(any(same_establishment == "y"), 1, 0), 
              .groups = "keep") %>% 
    ungroup()
  
  num_matched_estabs <- sum(num_matched_estabs$matched)
  
  num_matched_pats <- df %>% 
    group_by(patent_number) %>% 
    summarize(matched = ifelse(any(same_establishment == "y"), 1, 0), 
              .groups = "keep") %>% 
    ungroup()
  
  num_matched_pats <- sum(num_matched_pats$matched)
  
  data.frame("num_estabs" = num_estabs, 
             "num_matched_estabs" = num_matched_estabs,
             "num_pats" = num_pats,
             "num_matched_pats" = num_matched_pats)
}
```

```{r}
count_estabs_pats(dearborn_in)
```

```{r}
count_estabs_pats(springfield_oh)
```

# Linking Sheet Code

Make a linking sheet. Note: this is prototype code to scale these sheets up too!

```{r, eval = FALSE}
fips_code <- "39023"
year <- "1870"
linking_sheet_folder <- "~/Dropbox/Work/UChicago/RA/RA Hornbeck/CMF Patents/Terence Working Folders/output/pilot_linking_sheets/"
linking_sheet_path <- paste0(linking_sheet_folder, "fips", fips_code, "_", year, ".xlsx")

cmf_1870_county <- cmf_1870 %>%
  mutate(fips = as.character(fips),
         fips = str_pad(fips, width = 5, side = "left", pad = "0")) %>% 
  filter(fips == fips_code) %>% 
  select(firm_name, post_office, industry_raw1, machine_description1,
         materials_kinds_raw1, production_kinds_raw1, file_name, firm_number) %>% 
  rename(establishment_name = firm_name,
         industry = industry_raw1,
         machines = machine_description1,
         materials = materials_kinds_raw1,
         products = production_kinds_raw1)

patents_county <- patents %>% 
  filter(inventor_fips == fips_code|assignee_fips == fips_code) %>% 
  mutate(county_type = ifelse(inventor_fips == fips_code, 
                                       "inventor", "assignee"))

# Make linking sheet, get Jaro-Winkler, sort
linking_sheet <- merge(cmf_1870_county, patents_county, all = TRUE) %>% 
  mutate(same_establishment = "",
         jaro_winkler = stringdist(establishment_name, inventor_name,
                                   method = "jw", p = 0.1)) %>% 
  arrange(establishment_name, file_name, firm_number, jaro_winkler,
          inventor_name, assignee_name, patent_number) %>% 
  select(same_establishment, 
         establishment_name, inventor_name, assignee_name, # name variables
         county_type,
         industry, nber_subclass_name, nber_name, # industry related variables
         machines, materials, products, # industry related variables 2
         post_office, inventor_city, assignee_city, # finer location variables
         file_name, firm_number, patent_number, disposal_date) %>% # identifiers 
  mutate(across(.cols = everything(), as.character))

# Sheet formatting
# Replace all NAs with blanks
linking_sheet[is.na(linking_sheet)] <- ""

# Insert blank row between each CMF establishment block
linking_sheet %<>%
  group_by(establishment_name, file_name, firm_number) %>%
  do(add_row(.)) %>% 
  ungroup()
blank_indices <- which(rowSums(is.na(linking_sheet)) == ncol(linking_sheet))
blanks <- matrix(data = "", nrow = length(blank_indices), ncol = ncol(linking_sheet))
linking_sheet[blank_indices, ] <- blanks

# Make Excel sheet
write_formatted_excel <- function(df, output_path, 
                                  has_blanks = FALSE, blank_indices = NULL){
  # Create an Excel sheet
  excel <- createWorkbook()
  addWorksheet(excel, "Sheet 1")
  writeData(excel, sheet = "Sheet 1", df)
  
  # Color blank rows differently
  if (has_blanks){
    blank_fill <- createStyle(fgFill = "gray95")
    for (index in blank_indices){
      addStyle(excel, "Sheet 1", blank_fill, 
               rows = index + 1, cols = 1:ncol(df))
    }
  }
  
  # Color CMF, inventor and assignee columns differently
  index_inventor <- str_which(colnames(df), pattern = "inventor|nber")
  cell_fill <- createStyle(fgFill = "gray80", border = "TopBottomLeftRight", 
                           borderColour = "darkgray")
  for (index in index_inventor){
    addStyle(excel, "Sheet 1", cell_fill, 
             rows = 1:(nrow(df) + 1), cols = index)
  }
  
  index_assignee <- str_which(colnames(df), pattern = "assignee")
  cell_fill <- createStyle(fgFill = "wheat2", border = "TopBottomLeftRight", 
                           borderColour = "darkgray")
  for (index in index_assignee){
    addStyle(excel, "Sheet 1", cell_fill, 
             rows = 1:(nrow(df) + 1), cols = index)
  }
  
  # Add borders
  border_top <- createStyle(border = "Bottom", borderStyle = "medium")
  border_top_2 <- createStyle(border = "BottomRight", borderStyle = "medium")
  border_middle <- createStyle(border = "LeftRight", borderStyle = "medium")
  addStyle(excel, "Sheet 1", border_top, 
           rows = 1, cols = 1:ncol(df))
  addStyle(excel, "Sheet 1", border_top_2, rows = 1, cols = 1)
  addStyle(excel, "Sheet 1", border_middle, 
           rows = 2:(nrow(df) + 1), cols = 1)
  
  # Fix blank row borders
  if (has_blanks){
    blank_fill_2 <- createStyle(fgFill = "gray95", border = "LeftRight", 
                                borderStyle = "medium")
    addStyle(excel, "Sheet 1", blank_fill_2, 
             rows = (blank_indices + 1), cols = 1)
  }
  
  # Freze first row
  freezePane(excel, "Sheet 1", firstRow = TRUE)
  
  # Adjust column widths
  width_vec <- apply(df, 2, function(x){
    max(nchar(as.character(x)) + 1, na.rm = TRUE)
  })
  width_vec_header <- nchar(colnames(df)) + 1
  width_vec_max <- pmax(width_vec, width_vec_header)
  setColWidths(excel, "Sheet 1", 
               cols = 1:ncol(df), widths = width_vec_max)
  
  # Export
  saveWorkbook(excel, file = output_path, overwrite = TRUE)
}
  
# write_formatted_excel(df = linking_sheet, output_path = linking_sheet_path,
#                      has_blanks = TRUE, blank_indices = blank_indices)
```







